Medicare Provider Charge Data: Inpatient
========================================

Last update by Benjamin Chan (<benjamin.ks.chan@gmail.com>) on `r paste(Sys.time())` using `r R.version.string`.

Analyze CMS Medicare Provider Charge Data: Inpatient public use dataset. The data is documented and can be downloaded at the Medicare Provider Charge Data: Inpatient [website](http://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Inpatient.html).

>The data provided here include hospital-specific charges for the more than 3,000 U.S. hospitals that receive Medicare Inpatient Prospective Payment System (IPPS) payments for the top 100 most frequently billed discharges, paid under Medicare based on a rate per discharge using the Medicare Severity Diagnosis Related Group (MS-DRG) for Fiscal Year (FY) 2011. These DRGs represent almost 7 million discharges or 60 percent of total Medicare IPPS discharges.

>Hospitals determine what they will charge for items and services provided to patients and these charges are the amount the hospital bills for an item or service. The Total Payment amount includes the MS-DRG amount, bill total per diem, beneficiary primary payer claim payment amount, beneficiary Part A coinsurance amount, beneficiary deductible amount, beneficiary blood deducible amount and DRG outlier amount.

>For these DRGs, average charges and average Medicare payments are calculated at the individual hospital level. Users will be able to make comparisons between the amount charged by individual hospitals within local markets, and nationwide, for services that might be furnished in connection with a particular inpatient stay.

--------------------------------------------------------------------------------

Reading and manipulating the data
---------------------------------

Load the required libraries.
```{r LoadLibraries, results='hide'}
require(RCurl, quietly=TRUE)
require(xtable, quietly=TRUE)
require(ggplot2, quietly=TRUE)
require(scales, quietly=TRUE)
require(RColorBrewer, quietly=TRUE)
```

RCurl with https is being tempermental. Copy the dataset to the local folder and `read.table` from there. Read the entire dataset. 
```{r ReadData}
# url <- getURL("https://dl.dropboxusercontent.com/u/386956/Medicare_Provider_Charge_Inpatient_DRG100_FY2011.csv")
# df <- read.csv(textConnection(url), header=TRUE, sep=",")
df <- read.csv("Medicare_Provider_Charge_Inpatient_DRG100_FY2011.csv", header=TRUE, sep=",")
```

The raw data has `r nrow(df)` rows. Field names and the first few rows are below.
```{r PeekAtRawDataset}
names(df)
head(df)
```

Also read a MDC-DRG mapping table. This table comes from [CMS](http://www.cms.gov/Medicare/Medicare-Fee-for-Service-Payment/AcuteInpatientPPS/FY-2013-IPPS-Final-Rule-Home-Page-Items/FY2013-Final-Rule-Tables.html), [Table 5](http://www.cms.gov/Medicare/Medicare-Fee-for-Service-Payment/AcuteInpatientPPS/Downloads/FY_13_FR_Table_5.zip). Finally, create an MDC label table.
```{r ReadMappingTables, tidy=FALSE}
drg <- read.table("CMS-1588-F TABLE 5.txt", skip=2, nrows=751, sep="\t", col.names=c("drg", "postacute", "specialpay", "mdc", "type", "drglab", "drgwgt", "meanLOSg", "meanLOGa"))
mdc <- matrix(
  ncol=2,
  byrow=TRUE,
  data=c(
"PRE", "Pre-MDC",
"01" , "Nervous System",
"02" , "Eye",
"03" , "Ear, Nose, Mouth And Throat",
"04" , "Respiratory System",
"05" , "Circulatory System",
"06" , "Digestive System",
"07" , "Hepatobiliary System And Pancreas",
"08" , "Musculoskeletal System And Connective Tissue",
"09" , "Skin, Subcutaneous Tissue And Breast",
"10" , "Endocrine, Nutritional And Metabolic System",
"11" , "Kidney And Urinary Tract",
"12" , "Male Reproductive System",
"13" , "Female Reproductive System",
"14" , "Pregnancy, Childbirth And Puerperium",
"15" , "Newborn And Other Neonates (Perinatal Period)",
"16" , "Blood and Blood Forming Organs and Immunological Disorders",
"17" , "Myeloproliferative DDs (Poorly Differentiated Neoplasms)",
"18" , "Infectious and Parasitic DDs",
"19" , "Mental Diseases and Disorders",
"20" , "Alcohol/Drug Use or Induced Mental Disorders",
"21" , "Injuries, Poison And Toxic Effect of Drugs",
"22" , "Burns",
"23" , "Factors Influencing Health Status",
"24" , "Multiple Significant Trauma",
"25" , "Human Immunodeficiency Virus Infection"
    ),
  )
mdc <- data.frame(mdc)
names(mdc) <- c("mdc", "mdclab")
drg <- merge(drg, mdc, by=c("mdc"))
```

Create some new fields. 
* `DRGnum` is a numeric vector of the MS-DRG number without the description
* `DRGchar` is a character vector of the MS-DRG number without the description
* `isOregon` is a logical vector indicating if the row is from provider in Oregon
* `isPDXmetro` is a logical vector indicating if the row is from a provider in the Portland metro hospital referral region
* `isOHSU` is a logical vector indicating if the row is from an OHSU provider
```{r CreateVariables}
df$DRGnum <- as.numeric(substr(df$DRG.Definition, 1, 3))
df$DRGchar <- substr(df$DRG.Definition, 1, 3)
df$isOregon <- df$Provider.State == "OR"
df$isPDXmetro <- grepl("^OR - Portland", df$Hospital.Referral.Region.Description)
df$isOHSU <- grepl("^OHSU", df$Provider.Name)
```

Now that we have DRG in numeric form, merge the dataset to the MDC-DRG lookup.
```{r MergeMDCDRG}
df <- merge(df, drg, by.x=c("DRGnum"), by.y=c("drg"))
```

Get vector of MS-DRGs and MDCs that OHSU has data for.
```{r OHSU-DRGs}
drgOHSU <- as.vector(df$DRGnum[df$isOHSU])
mdcOHSU <- as.vector(unique(df$mdc[df$isOHSU]))
```

Create a subset of rows with just the DRGs in vector `drgOHSU` and non-missing DRG type.
```{r CreateSubset}
dfSubset <- subset(df, DRGnum %in% drgOHSU & !is.na(type))
```

The subset data has `r nrow(dfSubset)` rows. Field names and the first few rows are below.
```{r PeekAtAnalyticDataset}
names(dfSubset)
head(dfSubset)
```

Plot the data
-------------

Create function for plotting average covered charges.
```{r DefinePlotFunction, tidy=FALSE}
DistnPlot <- function(d, t) {
  ggplot(d, aes(x=DRGchar, y=Average.Covered.Charges)) +
#     geom_boxplot(alpha=1/2, fill="grey", outlier.size=0) +
    geom_violin(alpha=1/2, fill="grey") +
    geom_jitter(data=d[d$isPDXmetro,], alpha=1/2, na.rm=TRUE) + 
    geom_point(data=d[d$isOHSU,], alpha=2/3, color=rgb(0, 73, 144, max=255), size=5) + 
    facet_wrap(~ type, ncol=1, scales="free") +
    scale_y_log10(breaks=c(1E4, 2E4, 4E4, 8E4, 16E4, 32E4), labels=c("$10", "$20", "$40", "$80", "$160", "$320")) +
    labs(title=t, x="DRG", y="Average Covered Charges ($1,000s)") +
    theme_bw() +
    theme(legend.position="bottom", panel.grid.major.x=element_blank(), panel.grid.major.y=element_line(color="grey"))
}
```

Plot each MDC. Facet by DRG type.
* Blue dot is OHSU
* Grey dots are Portland metro hospitals
* Violins represent the nationwide distribution
```{r AverageCoveredCharges}
for (i in 1:length(mdcOHSU)) {
  j <- mdcOHSU[i]
  d <- subset(dfSubset, mdc == j)
  t <- paste("MDC", mdc$mdc[mdc == j], mdc$mdclab[mdc == j])
  show(DistnPlot(d, t))
}
```