Create MDC subsets.
```{r CreateMDCSubsets, eval=FALSE}
dfMDC01S <- subset(df, 020 <= DRGnum & DRGnum <= 042)
dfMDC01M <- subset(df, 052 <= DRGnum & DRGnum <= 103)
dfMDC02S <- subset(df, 113 <= DRGnum & DRGnum <= 117)
dfMDC02M <- subset(df, 121 <= DRGnum & DRGnum <= 125)
dfMDC03S <- subset(df, 129 <= DRGnum & DRGnum <= 139)
dfMDC03M <- subset(df, 146 <= DRGnum & DRGnum <= 159)
dfMDC04S <- subset(df, 163 <= DRGnum & DRGnum <= 168)
dfMDC04M <- subset(df, 175 <= DRGnum & DRGnum <= 206)
dfMDC05S <- subset(df, 215 <= DRGnum & DRGnum <= 265)
dfMDC05M <- subset(df, 280 <= DRGnum & DRGnum <= 316)
dfMDC06S <- subset(df, 326 <= DRGnum & DRGnum <= 358)
dfMDC06M <- subset(df, 368 <= DRGnum & DRGnum <= 395)
dfMDC07S <- subset(df, 405 <= DRGnum & DRGnum <= 425)
dfMDC07M <- subset(df, 432 <= DRGnum & DRGnum <= 446)
dfMDC08S <- subset(df, 453 <= DRGnum & DRGnum <= 517)
dfMDC08M <- subset(df, 533 <= DRGnum & DRGnum <= 566)
dfMDC09S <- subset(df, 570 <= DRGnum & DRGnum <= 585)
dfMDC09M <- subset(df, 592 <= DRGnum & DRGnum <= 607)
dfMDC10S <- subset(df, 614 <= DRGnum & DRGnum <= 630)
dfMDC10M <- subset(df, 637 <= DRGnum & DRGnum <= 645)
dfMDC11S <- subset(df, 652 <= DRGnum & DRGnum <= 675)
dfMDC11M <- subset(df, 682 <= DRGnum & DRGnum <= 700)
dfMDC12S <- subset(df, 707 <= DRGnum & DRGnum <= 718)
dfMDC12M <- subset(df, 722 <= DRGnum & DRGnum <= 730)
dfMDC13S <- subset(df, 734 <= DRGnum & DRGnum <= 750)
dfMDC13M <- subset(df, 754 <= DRGnum & DRGnum <= 761)
dfMDC14S <- subset(df, 765 <= DRGnum & DRGnum <= 770)
dfMDC14M <- subset(df, 774 <= DRGnum & DRGnum <= 782)
dfMDC15M <- subset(df, 789 <= DRGnum & DRGnum <= 795)
dfMDC16S <- subset(df, 799 <= DRGnum & DRGnum <= 804)
dfMDC16M <- subset(df, 808 <= DRGnum & DRGnum <= 816)
dfMDC17S <- subset(df, 820 <= DRGnum & DRGnum <= 830)
dfMDC17M <- subset(df, 834 <= DRGnum & DRGnum <= 849)
dfMDC18S <- subset(df, 856 <= DRGnum & DRGnum <= 855)
dfMDC18M <- subset(df, 862 <= DRGnum & DRGnum <= 872)
dfMDC19S <- subset(df, 876 == DRGnum)
dfMDC19M <- subset(df, 880 <= DRGnum & DRGnum <= 887)
dfMDC20M <- subset(df, 894 <= DRGnum & DRGnum <= 897)
dfMDC21S <- subset(df, 901 <= DRGnum & DRGnum <= 909)
dfMDC21M <- subset(df, 913 <= DRGnum & DRGnum <= 923)
dfMDC22S <- subset(df, 927 <= DRGnum & DRGnum <= 934)
dfMDC22S <- subset(df, 935 == DRGnum)
dfMDC23S <- subset(df, 939 <= DRGnum & DRGnum <= 941)
dfMDC23M <- subset(df, 945 <= DRGnum & DRGnum <= 951)
dfMDC24S <- subset(df, 955 <= DRGnum & DRGnum <= 959)
dfMDC24M <- subset(df, 963 <= DRGnum & DRGnum <= 965)
dfMDC25S <- subset(df, 969 <= DRGnum & DRGnum <= 970)
dfMDC25M <- subset(df, 974 <= DRGnum & DRGnum <= 977)
```
According to [findacode.com](http://www.findacode.com/code-set.php?set=DRG&mdc=04), DRGs 207 and 208 (respiratory system with ventilator support) are grouped as medical DRGs under MDC 04 (diseases & disorders of the respiratory system). This doesn't seem right. I'm going to exclude these DRGs until further clarification.
DRGs 998 (principal diagnosis invalid as discharge diagnosis) and 999 (ungroupable) are also excluded.


Surgical DRGs
-------------
Violin plots for surgical DRGs by MDC. Code chunks that are not evaluated do not have data. *NEED TO TWEAK THE FUNCTION TO CHECK THIS AND RETURN NULL*
```{r DistnPlotMDC01S}
DistnPlot(dfMDC01S, "MDC 01 Diseases & disorders of the nervous system")
```
```{r DistnPlotMDC02S, eval=FALSE}
DistnPlot(dfMDC02S, "MDC 02 Diseases & disorders of the eye")
```
```{r DistnPlotMDC03S, eval=FALSE}
DistnPlot(dfMDC03S, "MDC 03 Diseases & disorders of the ear, nose, mouth & throat")
```
```{r DistnPlotMDC04S, eval=FALSE}
DistnPlot(dfMDC04S, "MDC 04 Diseases & disorders of the respiratory system")
```
```{r DistnPlotMDC05S}
DistnPlot(dfMDC05S, "MDC 05 Diseases & disorders of the circulatory system")
```
```{r DistnPlotMDC06S}
DistnPlot(dfMDC06S, "MDC 06 Diseases & disorders of the digestive system")
```
```{r DistnPlotMDC07S}
DistnPlot(dfMDC07S, "MDC 07 Diseases & disorders of the hepatobiliary system & pancreas")
```
```{r DistnPlotMDC08S}
DistnPlot(dfMDC08S, "MDC 08 Diseases & disorders of the musculoskeletal system & connective tissue")
```
```{r DistnPlotMDC09S, eval=FALSE}
DistnPlot(dfMDC09S, "MDC 09 Diseases & disorders of the skin, subcutaneous tissue & breast")
```
```{r DistnPlotMDC10S, eval=FALSE}
DistnPlot(dfMDC10S, "MDC 10 Endocrine, nutritional & metabolic diseases & disorders")
```
```{r DistnPlotMDC11S, eval=FALSE}
DistnPlot(dfMDC11S, "MDC 11 Diseases & disorders of the kidney & urinary tract")
```
```{r DistnPlotMDC12S, eval=FALSE}
DistnPlot(dfMDC12S, "MDC 12 Diseases & disorders of the male reproductive system")
```
```{r DistnPlotMDC13S, eval=FALSE}
DistnPlot(dfMDC13S, "MDC 13 Diseases & disorders of the female reproductive system")
```
```{r DistnPlotMDC14S, eval=FALSE}
DistnPlot(dfMDC14S, "MDC 14 Pregnancy, childbirth & the puerperium")
```
```{r DistnPlotMDC16S, eval=FALSE}
DistnPlot(dfMDC16S, "MDC 16 Disease & disorders of blood, blood forming organs, immunologic disorders")
```
```{r DistnPlotMDC17S, eval=FALSE}
DistnPlot(dfMDC17S, "MDC 17 Myeloproliferative diseases & disorders, poorly differentiated neoplasm")
```
```{r DistnPlotMDC18S, eval=FALSE}
DistnPlot(dfMDC18S, "MDC 18 Infectious & parasitic diseases, systemic or unspecified sites")
```
```{r DistnPlotMDC19S, eval=FALSE}
DistnPlot(dfMDC19S, "MDC 19 Mental diseases & disorders")
```
```{r DistnPlotMDC20S, eval=FALSE}
DistnPlot(dfMDC20S, "MDC 20 Alcohol/drug use & alcohol/drug induced organic mental disorders")
```
```{r DistnPlotMDC21S, eval=FALSE}
DistnPlot(dfMDC21S, "MDC 21 Injuries, poisonings & toxic effects of drugs")
```
```{r DistnPlotMDC22S, eval=FALSE}
DistnPlot(dfMDC22S, "MDC 22 Burns")
```
```{r DistnPlotMDC23S, eval=FALSE}
DistnPlot(dfMDC23S, "MDC 23 Factors influencing hlth stat & othr contacts with hlth servcs")
```
```{r DistnPlotMDC24S, eval=FALSE}
DistnPlot(dfMDC24S, "MDC 24 Multiple significant trauma")
```
```{r DistnPlotMDC25S, eval=FALSE}
DistnPlot(dfMDC25S, "MDC 25 Human immunodeficiency virus infections")
```

Medical DRGs
-------------
Violin plots for medical DRGs by MDC. Code chunks that are not evaluated do not have data.
```{r DistnPlotMDC01M}
DistnPlot(dfMDC01M, "MDC 01 Diseases & disorders of the nervous system")
```
```{r DistnPlotMDC02M, eval=FALSE}
DistnPlot(dfMDC02M, "MDC 02 Diseases & disorders of the eye")
```
```{r DistnPlotMDC03M}
DistnPlot(dfMDC03M, "MDC 03 Diseases & disorders of the ear, nose, mouth & throat")
```
```{r DistnPlotMDC04M}
DistnPlot(dfMDC04M, "MDC 04 Diseases & disorders of the respiratory system")
```
```{r DistnPlotMDC05M}
DistnPlot(dfMDC05M, "MDC 05 Diseases & disorders of the circulatory system")
```
```{r DistnPlotMDC06M}
DistnPlot(dfMDC06M, "MDC 06 Diseases & disorders of the digestive system")
```
```{r DistnPlotMDC07M}
DistnPlot(dfMDC07M, "MDC 07 Diseases & disorders of the hepatobiliary system & pancreas")
```
```{r DistnPlotMDC08M}
DistnPlot(dfMDC08M, "MDC 08 Diseases & disorders of the musculoskeletal system & connective tissue")
```
```{r DistnPlotMDC09M}
DistnPlot(dfMDC09M, "MDC 09 Diseases & disorders of the skin, subcutaneous tissue & breast")
```
```{r DistnPlotMDC10M}
DistnPlot(dfMDC10M, "MDC 10 Endocrine, nutritional & metabolic diseases & disorders")
```
```{r DistnPlotMDC11M}
DistnPlot(dfMDC11M, "MDC 11 Diseases & disorders of the kidney & urinary tract")
```
```{r DistnPlotMDC12M, eval=FALSE}
DistnPlot(dfMDC12M, "MDC 12 Diseases & disorders of the male reproductive system")
```
```{r DistnPlotMDC13M, eval=FALSE}
DistnPlot(dfMDC13M, "MDC 13 Diseases & disorders of the female reproductive system")
```
```{r DistnPlotMDC14M, eval=FALSE}
DistnPlot(dfMDC14M, "MDC 14 Pregnancy, childbirth & the puerperium")
```
```{r DistnPlotMDC15M, eval=FALSE}
DistnPlot(dfMDC15M, "MDC 15 Newborns & other neonates with condtn orig in perinatal period")
```
```{r DistnPlotMDC16M}
DistnPlot(dfMDC16M, "MDC 16 Disease & disorders of blood, blood forming organs, immunologic disorders")
```
```{r DistnPlotMDC17M, eval=FALSE}
DistnPlot(dfMDC17M, "MDC 17 Myeloproliferative diseases & disorders, poorly differentiated neoplasm")
```
```{r DistnPlotMDC18M}
DistnPlot(dfMDC18M, "MDC 18 Infectious & parasitic diseases, systemic or unspecified sites")
```
```{r DistnPlotMDC19M}
DistnPlot(dfMDC19M, "MDC 19 Mental diseases & disorders")
```
```{r DistnPlotMDC20M}
DistnPlot(dfMDC20M, "MDC 20 Alcohol/drug use & alcohol/drug induced organic mental disorders")
```
```{r DistnPlotMDC21M}
DistnPlot(dfMDC21M, "MDC 21 Injuries, poisonings & toxic effects of drugs")
```
```{r DistnPlotMDC22M, eval=FALSE}
DistnPlot(dfMDC22M, "MDC 22 Burns")
```
```{r DistnPlotMDC23M}
DistnPlot(dfMDC23M, "MDC 23 Factors influencing hlth stat & othr contacts with hlth servcs")
```
```{r DistnPlotMDC24M, eval=FALSE}
DistnPlot(dfMDC24M, "MDC 24 Multiple significant trauma")
```
```{r DistnPlotMDC25M, eval=FALSE}
DistnPlot(dfMDC25M, "MDC 25 Human immunodeficiency virus infections")
```
